---
title: "Handling categorical data in R"
output: html_notebook
---


# R

## Packages

```{r}
library(tidyverse)
library(forcats)
```

## factors / characters

We use `base::factor` function to create factor 

```{r}
data <- 1:10
data_f <- as.factor(data)
data_f
```

Using `base::levels` we get levels

```{r}
levels(data_f)
```

Now, create a factor with levels and labels defined

```{r}
f1 <- factor(x = c(1,2,3,1,2,3,-9), 
             levels = c(1,2,3,-9, 0),
             labels = c("Yes", "No", "Don't know", "N/A", "NULL"),
             ordered = FALSE)
f1
```

Note that `as.numeric` replaces original data!!! (no -9 values)

```{r}
as.numeric(f1)
```

Function `as.character` converts replaces numbers with labels.

```{r}
as.character(f1)
```

Using table

```{r}
table(f1)
```

To drop levels with 0 observations (not recommended for initial analysis) we can use `table(., exclude = "level")`

```{r}
table(f1, exclude = "NULL")
```

or `base::droplevels` function

```{r}
table(droplevels(f1))
```

or `xtabs(., drop.unused.levels = TRUE)`
```{r}
xtabs(~f1, drop.unused.levels = TRUE)
```

We can change the reference class (i.e. first level)  with `relevel` function.

```{r}
relevel(f1, ref = "No")
```

In R we can create an ordered factor (for ordered categorical data). 

```{r}
f2 <- factor(x = mtcars$cyl,
             levels = c(4, 6, 8),
             labels = c("Small", "Medium", "Big"),
             ordered = TRUE)

f2
```

Note `<` symbols in `levels`.


## other packages

### haven

```{r}
x <- labelled(x = c(1, 2, 1, 2, 10, 9), labels = c(Unknown = 9, Refused = 10))
x
```

```{r}
as.numeric(x)
```

```{r}
as_factor(x)
```

### forcats

Main functions

+ `fct_reorder()`: Reordering a factor by another variable.
+ `fct_infreq()`: Reordering a factor by the frequency of values.
+ `fct_relevel()`: Changing the order of a factor by hand.
+ `fct_lump()`: Collapsing the least/most frequent values of a factor into “other”.

Other functions -- https://forcats.tidyverse.org/reference/index.html


```{r}
fct_inorder(f = f2, ordered = F) ## in orde of appearance
```

```{r}
fct_infreq(f = f2, ordered = F) ## by frequency
```

```{r}
fct_count(f2) ## frequency table (result: data.frame)
```

```{r}
fct_lump(f2, n = 1, other_level = "other levels") ## aggregates based frequency
```

```{r}
fct_collapse(f2,
             other  = c("Big", "Medium"))

```

About missing levels

```{r}
f1 <- factor(c("a", "a", NA, NA, "a", "b", NA, "c", "a", "c", "b"))
table(f1)

f2 <- fct_explicit_na(f1, na_level = "Missing")
table(f2)
```


### Further reading

-- haven::labelled
-- dplyr::recode
-- expss -- https://cran.r-project.org/web/packages/expss/vignettes/labels-support.html



