---
title: "Fitting distributions"
output: 
  html_notebook: 
    toc: yes
---

# Fitting in R

## Packages

```{r}
library(vcdExtra)
library(fitdistrplus)
library(haven) ## read sas file
```


Generate variable from Poisson distribution

```{r}
set.seed(123)
x <- rpois(10000,10)
barplot(table(x))
```

### Fitting distributions using vcdExtra

Verify whether given variable follows binomial, negative binomial or poisson distribution

```{r}
m1 <- goodfit(x, "binomial")
m2 <- goodfit(x, "nbinomial")
m3 <- goodfit(x, "poisson")
summary(m1)
summary(m2)
summary(m3)
plot(m1)
plot(m2)
plot(m3)
```

```{r}
m1$par
m2$par
m3$par
```

### Fitting distributions using MASS::fitdistr

```{r}
fitdistr(x, "Poisson")
```

```{r}
fitdistr(x, "negative binomial")
```

### Fitting distributions using fitdistrplus::fitdist


```{r}
res1 <- fitdist(x, distr = "pois")
summary(res1)
```

```{r}
res2 <- fitdist(x, distr = "nbinom")
```

```{r}
fit1 <- fitdist(x, distr = "binom", fix.arg = list(size = max(x)), start = list(prob = 0.3))
```

## Practical example from real data

```{r}
bbgd <- read_sas("../data-raw/baza2005.sas7bdat")
```

```{r}
barplot(table(bbgd$LOS_LT14))
```

What kind of distribution it is? 

Is it poisson?

```{r}
dist_po <- goodfit(bbgd$LOS_LT14, "poisson")
```

```{r}
plot(dist_po)
```

```{r}
summary(dist_po)
```

Is it negative binomial?

```{r}
dis_nb <- goodfit(bbgd$LOS_LT14, "nbinom")
```

```{r}
summary(dis_nb)
```

```{r}
plot(dis_nb)
```

```{r}
dis_nb <- goodfit(bbgd$LOS_LT14, "binom")
```

```{r}
summary(dis_nb)
```

