---
title: "Fitting distributions"
output: 
  html_notebook: 
    toc: yes
---

# Fitting in R

## Packages

```{r}
library(vcdExtra)
library(fitdistrplus)
library(haven) ## read sas file
```


Generate variable from Poisson distribution

```{r}
set.seed(123)
x <- rpois(10000,10)
mean(x)
barplot(table(x))
```

### Fitting distributions using maximum likelihood

In case of when vector of parameter contains only one value use stats::optim with method Brent or stats::optimize

Using stats::optim

```{r}
pois_ll <- function(par, x) {
  ll <- -length(x) * par + sum(log(factorial(x))) + log(par) * sum(x)
  - 2 * ll
}

res <- optim(par = 0, fn = pois_ll, method = "Brent", lower = 0, upper = 15, x = x)
res
```

Using stats::optimize

```{r}
res2 <- optimize(f = pois_ll, interval = c(0,15), x = x)
res2
```

### Fitting distributions using vcdExtra

Verify whether given variable follows binomial, negative binomial or poisson distribution

```{r}
m1 <- goodfit(x, "binomial")
m2 <- goodfit(x, "nbinomial")
m3 <- goodfit(x, "poisson")
summary(m1)
summary(m2)
summary(m3)
plot(m1)
plot(m2)
plot(m3)
```

```{r}
m1$par
m2$par
m3$par
```

### Fitting distributions using MASS::fitdistr

```{r}
fitdistr(x, "Poisson")
```

```{r}
fitdistr(x, "negative binomial")
```

### Fitting distributions using fitdistrplus::fitdist


```{r}
res1 <- fitdist(x, distr = "pois")
summary(res1)
```

```{r}
res2 <- fitdist(x, distr = "nbinom")
```

```{r}
fit1 <- fitdist(x, distr = "binom", fix.arg = list(size = max(x)), start = list(prob = 0.3))
```

## Practical example from real data

```{r}
bbgd <- read_sas("../data-raw/baza2005.sas7bdat")
bbgd
```

Let's focus on variable informing how many children under 14 years are in households (variable LOS_LT14)


```{r}
barplot(table(bbgd$LOS_LT14))
```

What kind of distribution it is? 

Is it poisson?

```{r}
dist_po <- goodfit(bbgd$LOS_LT14, "poisson")
```

```{r}
plot(dist_po)
```

```{r}
summary(dist_po)
```

Is it negative binomial?

```{r}
dis_nb <- goodfit(bbgd$LOS_LT14, "nbinom")
```

```{r}
summary(dis_nb)
```

```{r}
plot(dis_nb)
```

```{r}
dis_nb <- goodfit(bbgd$LOS_LT14, "binom")
```

```{r}
summary(dis_nb)
```

# Fitting in Julia

```{julia, engine.path = "/Applications/Julia-1.0.app/Contents/Resources/julia/bin/julia"}
using Distributions ## for distributions
using Random
Random.seed!(123);
x = rand(Poisson(10),10000);
m = fit_mle(Poisson,x);
print(m)
```

# Fitting in Python

sciPy can only fit continuous distributions. 

