---
title: "R Notebook"
output: html_notebook
---

We will start with generating random numbers

```{r}
set.seed(2019)
var1 <- rbinom(n = 1000, ## numer of obs / length of vector
               size = 1, ## number of trials
               prob = 0.2)

head(var1)
table(var1)
```

generate numbers for Poisson distribution


```{r}
set.seed(123)
var2 <- rpois(n = 1000,
              lambda = 10)
table(var2)
barplot(table(var2))
```

Now, we generate varibles from multinomial distribution

```{r}
set.seed(2016)
var3 <- rmultinom(n = 10,
                  size = 100,
                  prob = c(0.1,0.3,0.6))

head(var3)
```


During trening basketball player throws 10 times for 3 points. Probability of success in a single shot is 0.4. Let $X$ be random variable denoting number of accurate throw.

+ Find the distribution of $X$.


```{r}
ex1a <- dbinom(x = 0:10, size = 10, prob = 0.4)

data.frame(k = 0:10, p = ex1a)
```

+ Create a plot for the distribution of  $X$.
+ Find probability that the player throws exactly 6 times

```{r}
dbinom(x = 6, size = 10, prob = 0.4)
```

+ Find probability that the player throws at most 3 times


```{r}
pbinom(q = 3, size = 10, prob = 0.4) 
## dbinom(0,10,0.4)+dbinom(1,10,0.4)+dbinom(2,10,0.4)+dbinom(3,10,0.4)
```

```{r}
# F(X>=5) = 1 - F(X <= 4)
1 - pbinom(q = 4, size = 10, prob = 0.4)
```

Poisson distribution

A car salesman sales on average 5 cars. What is the probability that this dealer will sale at most 2 cars?


Calculate probability of selling 2 cars

```{r}
dpois(x = 2, lambda = 5)
```

```{r}
ppois(q = 2, lambda = 5)
```

A young mathematician writing his articles in the LaTeX program stated that he commits an average of 4 typos per page. We randomly select one of the many pages written by this mathematician. What is the probability that we will find at most 3 errors on it.

```{r}
ppois(q = 3, lambda = 4)
```

```{r}
1-ppois(q = 4, lambda = 4)
```

